Rust for Operating System Development: A Systematic Literature Review

### 2024-04-22 paper.tex
Updated search strings for Scopus database

In the ever-evolving landscape of software development, the reliability and security of computer
systems stands as a paramount concern for all parties involved. Modern software is constructed by
building ever more complex abstractions, one on top of the other. Thus, if we aim to have a secure
system we must start to peel back all the layers and tackle one of the fundamental abstractions in
computer science, the programming language. Programming languages that provided and enforce memory
safety eliminate whole classes of bugs such as buffer overflows, dangling pointers, and memory leaks
which have been implicated in a myriad of security vulnerabilities and system crashes.

@@

Before we start researching we must first confirm the need for a SLR. The recent report released by
ONCD~\cite{United_States_Gov2024-pp} has conveniently done this job for us by compiling a report
detailing the need for research in the domain of memory safety. While the ONCD report detailed a two
pronged approach for this SLR we will be focusing on a memory safe programming language,
specifically Rust. While there are many memory safe programming languages available for software
developers to use Rust is the only language that has support in kernel
space~\cite{The_kernel_development_community_undated-iw} thus is a primary candidate to replace the
aging C programming language.

@@

We employed a robust multi-step search strategy across three digital databases, ACM Digital Library,
IEEE Xplore, and Google Scholar in order to find all the current research that has been done. These
databases cover a wide range of current software engineering and computer science literature. We
leveraged the advanced search features of both the ACM Digital Library and IEEE Xplore to search the
titles and abstracts for keywords using boolean search operators. The ACM Digital Library and IEEE
Xplore formed the core of our keyword search and we used Google Scholar for backward and forward
snowballing.

@@

We first searched abstracts in the ACM Digital Library and metadata in IEEE Xplore with the keywords
listed in table~\ref{tab:keywords} between the years 2019~\cite{Gaynor2019-dk} and 2024. We then
used Google Scholar for forward and backward snowballing.

\begin{table}
\begin{tabular}{|| p{2cm}| p{4cm} | p{1cm} ||}
 \hline
 Database & Query & Results \\
 \hline\hline
 ACM  & Abstract:("operating system" kernel linux) AND Abstract:(rust) & 22 \\
 IEEE & ("All Metadata":"operating system" AND "All Metadata":rust) OR ("All Metadata":kernel AND
 "All Metadata":rust) OR ("All Metadata":linux AND "All Metadata":rust) & 37 \\
 \hline
\end{tabular}
\caption{Search Queries used for each Database.}
\label{tab:keywords}
\end{table}

>>>>

In the ever-evolving landscape of software development, the reliability and security of computer
systems stands as a paramount concern for all parties involved. Modern software is constructed by
building ever more complex abstractions, one on top of the other. Thus, if we aim to have a secure
system we must start to peel back all the layers and tackle one of the fundamental abstractions in
computer science, the programming language. Programming languages that provide and enforce memory
safety eliminate whole classes of bugs such as buffer overflows, dangling pointers, and memory leaks
which have been implicated in a myriad of security vulnerabilities and system crashes.

@@

Before we start researching we must first confirm the need for a SLR. The recent report released by
ONCD~\cite{United_States_Gov2024-pp} has conveniently done this job for us by compiling a report
detailing the need for research in the domain of memory safety. While the ONCD report detailed a two
pronged approach, for this SLR we will be focusing on a memory safe programming language,
specifically Rust. While there are many memory safe programming languages available for software
developers to use Rust is the only language that has support in kernel
space~\cite{The_kernel_development_community_undated-iw} thus is a primary candidate to replace the
aging C programming language.

@@

We employed a robust multi-step search strategy across three digital databases, ACM Digital Library,
IEEE Xplore, and Scopus in order to find all the current research that has been done. These
databases cover a wide range of current software engineering and computer science literature. We
leveraged the advanced search features of all three databases to search the titles and abstracts for
keywords using boolean search operators.

@@

We search all three databases with the keywords listed in table~\ref{tab:keywords} between January
1, 2019~\cite{Gaynor2019-dk} and April 1, 2024. While both ACM and IEEE had mutually exclusive
results Scopus had 44 duplicates that needed to be removed. We then used Google Scholar for forward
and backward snowballing.

\begin{table}
\begin{tabular}{|| p{2cm}| p{4cm} | p{1cm} ||}
 \hline
 Database & Query & Results \\
 \hline\hline
 ACM  & (Abstract:("operating system" kernel linux) AND Abstract:(rust) )OR (Title:("operating
 system" kernel linux) AND Title:(rust)) & 25 \\
 IEEE & ("All Metadata":"operating system" AND "All Metadata":rust) OR ("All Metadata":kernel AND
 "All Metadata":rust) OR ("All Metadata":linux AND "All Metadata":rust) & 39 \\
 Scopus & TITLE-ABS-KEY ( ( "operating system" OR kernel OR linux ) AND rust ) AND PUBYEAR > 2018
 AND PUBYEAR < 2025 AND ( LIMIT-TO ( SUBJAREA , "COMP" ) ) AND ( LIMIT-TO ( LANGUAGE , "English" ) )
 & 87 \\
 \hline
\end{tabular}
\caption{Search Queries used for each Database.}
\label{tab:keywords}
\end{table}
